---
title: "Assignment 2"
author: " Vu Duc Thanh"
output: html_document
---


```{r echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
```

```{r}
data <- read.csv("gardasil.csv",header=T)
summary(data)
str(data)

```
```{r}
head(data)

```

```{r fig.width=5, fig.height=3}
data %>%
  ggplot(aes(x=Completed.within.12.months)) +
  geom_bar(stat="count",color = "black", fill = "gray") +
  xlab('Count') + ylab('Completed within 12 months') + ggtitle('Completed within 12 months')
```

```{r fig.width=5, fig.height=3}
data %>%
  ggplot(aes(Age, Completed.within.12.months)) + 
  geom_boxplot()+
  coord_flip()
```
```{r fig.height=15, fig.width=15}
library(GGally)
ggpairs(data)

```


```{r fig.width=12, fig.height=12,message=FALSE}
require(gridExtra)
p1<-data %>%
    ggplot(aes(x=AgeGroup,fill=Completed.within.12.months)) + 
        geom_bar(stat="count",color = "black",alpha = 0.65) +
        scale_fill_manual(values = c("#0072B2", "#009E73"))
p2<-data %>%
    ggplot(aes(x=Race,fill=Completed.within.12.months)) + 
        geom_bar(stat="count",color = "black",alpha = 0.65) +
        scale_fill_manual(values = c("#0072B2", "#009E73"))
#p3<-data %>%
#    ggplot(aes(x=Shots,fill=Completed.within.12.months)) + 
#        geom_histogram(binwidth = 0.5, color = "black",alpha = 0.65)+
#        scale_fill_manual(values = c("#0072B2", "#009E73"))
p3<-data %>%
    ggplot(aes(x=Insurance.type,fill=Completed.within.12.months)) + 
        geom_bar(stat="count",color = "black",alpha = 0.65) +
        scale_fill_manual(values = c("#0072B2", "#009E73"))
p4 <- data %>%
     ggplot(aes(x=Clinic,fill=Completed.within.12.months)) + 
        geom_bar(stat="count",color = "black",alpha = 0.65) +
        scale_fill_manual(values = c("#0072B2", "#009E73"))
p5 <- data %>%
    ggplot(aes(x=Location.type,fill=Completed.within.12.months)) + 
        geom_bar(stat="count",color = "black",alpha = 0.65) +
        scale_fill_manual(values = c("#0072B2", "#009E73"))
p6 <- data %>%
    ggplot(aes(x=Practice.type,fill=Completed.within.12.months)) + 
        geom_bar(stat="count",color = "black",alpha = 0.65)+
        scale_fill_manual(values = c("#0072B2", "#009E73"))
grid.arrange(p1,p2,p3,p4,p5,p6, ncol=2)
#plot(data,"Shots","Completed.within.12.months")
```

```{r}
chisq.test(data$AgeGroup,data$Completed.within.12.months, correct = F)
chisq.test(data$Race,data$Completed.within.12.months, correct = F)
chisq.test(data$Insurance.type,data$Completed.within.12.months, correct = F)
chisq.test(data$Clinic,data$Completed.within.12.months, correct = F)
chisq.test(data$Location.type,data$Completed.within.12.months, correct = F)
chisq.test(data$Practice.type,data$Completed.within.12.months, correct = F)

```
```{r}
chisq.test(data$AgeGroup,data$Completed.within.12.months, correct = F)$stdres
chisq.test(data$Race,data$Completed.within.12.months, correct = F)$stdres
chisq.test(data$Insurance.type,data$Completed.within.12.months, correct = F)$stdres
chisq.test(data$Clinic,data$Completed.within.12.months, correct = F)$stdres
chisq.test(data$Location.type,data$Completed.within.12.months, correct = F)$stdres
chisq.test(data$Practice.type,data$Completed.within.12.months, correct = F)$stdres

```
 

```{r}
cat(paste("black race completed the test",sum(data$Race=="black" &data$Completed.within.12.months=="yes")/sum(data$Race=="black")))

cat("\n")
cat(paste("white race completed the test",sum(data$Race=="white" &data$Completed.within.12.months=="yes")/sum(data$Race=="white")))

cat("\n")
cat(paste("Medical assistance completed the test",sum(data$Insurance.type=="medical assistance" &data$Completed.within.12.months=="yes")/sum(data$Insurance.type=="medical assistance")))

cat("\n")
cat(paste("Without medical assistance completed the test",sum(data$Insurance.type!="medical assistance" &data$Completed.within.12.months=="yes")/sum(data$Insurance.type!="medical assistance")))

cat("\n")
cat(paste("Clinic Bayview completed the test",sum(data$Clinic=="Bayview" &data$Completed.within.12.months=="yes")/sum(data$Clinic=="Bayview")))

cat("\n")
cat(paste("Clinic White Marsh completed the test",sum(data$Clinic=="White Marsh" &data$Completed.within.12.months=="yes")/sum(data$Clinic=="White Marsh")))


cat("\n")
cat(paste("People in suburban  completed the test",sum(data$Location.type=="suburban" &data$Completed.within.12.months=="yes")/sum(data$Location.type=="suburban")))

cat("\n")
cat(paste("People in urban  completed the test",sum(data$Location.type=="urban" &data$Completed.within.12.months=="yes")/sum(data$Location.type=="urban")))

```
```{r}
cat(paste("black race completed the test",sum(data$Race=="black" &data$Completed.within.12.months=="no")/sum(data$Race=="black")))

cat("\n")
cat(paste("white race completed the test",sum(data$Race=="white" &data$Completed.within.12.months=="no")/sum(data$Race=="white")))

cat("\n")
cat(paste("Medical assistance completed the test",sum(data$Insurance.type=="medical assistance" &data$Completed.within.12.months=="no")/sum(data$Insurance.type=="medical assistance")))

cat("\n")
cat(paste("Without medical assistance completed the test",sum(data$Insurance.type!="medical assistance" &data$Completed.within.12.months=="no")/sum(data$Insurance.type!="medical assistance")))

cat("\n")
cat(paste("Clinic Bayview completed the test",sum(data$Clinic=="Bayview" &data$Completed.within.12.months=="no")/sum(data$Clinic=="Bayview")))

cat("\n")
cat(paste("Clinic White Marsh completed the test",sum(data$Clinic=="White Marsh" &data$Completed.within.12.months=="no")/sum(data$Clinic=="White Marsh")))


cat("\n")
cat(paste("People in suburban  completed the test",sum(data$Location.type=="suburban" &data$Completed.within.12.months=="no")/sum(data$Location.type=="suburban")))

cat("\n")
cat(paste("People in urban  completed the test",sum(data$Location.type=="urban" &data$Completed.within.12.months=="no")/sum(data$Location.type=="urban")))



```

```{r}
surburban_no <- sum(data$Location.type=="suburban" & data$Completed.within.12.months=="no")
surburban_yes <- sum(data$Location.type=="suburban" & data$Completed.within.12.months=="yes")

urban_no <- sum(data$Location.type=="urban" & data$Completed.within.12.months=="no")
urban_yes <- sum(data$Location.type=="urban" & data$Completed.within.12.months=="yes")

sum_surburban <- sum(data$Location.type=="suburban")

sum_urban <- sum(data$Location.type=="urban")

prop.test(x = c(surburban_yes, urban_yes),
          n = c(sum_surburban, sum_urban),
          alternative = "two.sided",
          correct = FALSE)
```
```{r}
age_fisrt_no <- sum(data$AgeGroup=="11-17 years" & data$Completed.within.12.months=="no")
age_fisrt_yes <- sum(data$AgeGroup=="11-17 years" & data$Completed.within.12.months=="yes")

age_second_no <- sum(data$AgeGroup=="18-26 years" & data$Completed.within.12.months=="no")
age_second_yes <- sum(data$AgeGroup=="18-26 years" & data$Completed.within.12.months=="yes")

sum_age_fisrt <- sum(data$AgeGroup=="11-17 years")

sum_age_second <- sum(data$AgeGroup=="18-26 years")

prop.test(x = c(age_fisrt_yes, age_second_yes),
          n = c(sum_age_fisrt, sum_age_second),
          alternative = "greater",
          correct = FALSE)

```
Compare mean t-test
```{r}
yes_age<- data$Age[data$Completed.within.12.months=='yes']
no_age <- data$Age[data$Completed.within.12.months=='no']
t.test(yes_age,no_age)
```
Conclusion:

##Chi- square test

The residuals are approximately normal, so we could check the cells with values >3 and <-3, for example.

Results:

Black and white race are significant.

Medical assistance in insurance type is significant.

Clinic location at Bay view and white mash are significant.

Location type (surbuban and urban) data is significant.

Completed: 

           23% of black race completed the test vs 38% of white people

           20% of people completed medical assistance vs 36% of others
           
           26% of clinic Bayview vs 48% of clinic White Marsh
           
           36% of people in surburban vs 25% of people in urban
          
Not finished: 

          77% of black race  vs 62% of white people

          80% of people with medical assistance vs 64% of others
          
          74% of clinic Bayview vs 52% of clinic White Marsh
          
          64% of people in surburban vs 75% of people in urban
          
##Perform proportion test:

People in suburban are likely to completed compared to who live in urban area with p_value= 1.802e-05


People with the age between 11-17 tended to fished than people with age from 18-26 with p_value= 0.006472

##t-test

The mean of of age in both "yes" and "no" are equal. 